### 2019-04-25 report/chapter03.Rnw
presentation u.a.

<<echo=FALSE>>=
library(knitr)
opts_chunk$set(fig.path='figure/ch02_fig',
               echo=TRUE, message=FALSE,
               fig.width=8.1, fig.height=3.7,
               out.width='\\textwidth-3cm',
               message=FALSE, fig.align='center',
               background="gray98", tidy=FALSE, #tidy.opts=list(width.cutoff=60),
               cache=TRUE
)
options(width=74)
@

@@

\begin{figure}
<<echo=FALSE>>=
data %>% filter(total1 + total2 > 10 & total1 + total2 < 100) %>% 
  filter(outcome.measure.new == "Risk Ratio" | outcome.measure.new == "Odds Ratio" | outcome.measure.new == "Mean Difference" |
           outcome.measure.new == "Std. Mean Difference") %>% 
  group_by(outcome.measure.new) %>% 
  mutate(sample.size = total1 + total2, scaled.effect = abs(scale(effect, center = T, scale = T))) %>%
  group_by(sample.size, outcome.measure.new) %>% 
  summarize(median.effect = median(scaled.effect, na.rm = T)) %>% 
  ggplot(aes(x = sample.size, y = median.effect, group = outcome.measure.new)) + geom_point() + facet_wrap(~outcome.measure.new) + 
  theme_bw() + xlab("sample size") + ylab("median absolute normalized effect size")
@
\caption{Median of the absolute value of the normalized effect size plotted against the total sample size, separated for outcome measures.}
\label{effect.samplesize.separated}
\end{figure}

A second way to analyse meta-analyses is a cumulative meta-analysis that should reveal shifts in treatment effect sizes over time. This can again be done for the entire dataset, i.e. for all effect estimates and their ordering in time. It is important to scale the effect estimates here with respect to the estimates of the replication studies that are about the same subject, have the same outcome measure, etc. Also, only effect estimates that can be compared to other estimates before are included, i.e. only study results with one or more replica are included. Time needs to be scaled and normalized in order to compare multiple time-trends in effect size to each other and gain insights in the overall trend. This is done in Figure \ref{effects.overtime.separated}, separared for odds rati, risk ratio, and mean and standardized difference outcome measures. In order to reduce the spread over time of effect sizes, only studies between 1970 and 2019 were included, as well as studies with a minimal group size of 12 participants (each group).

\begin{figure}
<<echo=FALSE>>=
data.ext %>% filter(outcome.measure.new == "Mean Difference" | outcome.measure.new == "Std. Mean Difference" | 
                outcome.measure.new == "Risk Ratio" | outcome.measure.new == "Odds Ratio") %>% 
  filter(total1 > 11 & total2 > 11) %>% filter(!is.na(study.year) & study.year > 1970 & study.year < 2019) %>% 
  group_by(meta.id) %>% mutate(n = n()) %>% filter(n > 1) %>% 
  mutate(scaled.effect = abs(scale(effect)), scaled.time = scale(study.year)) %>% 
  ggplot(aes(x = scaled.time, y = scaled.effect)) + geom_point(size = .5, alpha = 0.2)  + facet_wrap(~outcome.measure.new) +
  geom_smooth() + theme_bw() +  ylab("absolute scaled effect")  + ylab("scaled time")
@
\caption{Absolute values of scaled effect sizes over scaled time, separated for outcome measures.}
\label{effects.overtime.separated}
\end{figure}

@@

\vspace{0mm}
Test performance depends on the sample size, despite having restricted sample size to a minimum of 10 studies. The p-values of the Thompson and Sharp tests are shown with respect to the sample size of the meta-analysis in Figure \ref{pvalues.samplesize}. %(Suggests that 10 is likely too small)
In the case of binary outcomes, the arcsine variance stabilizing function has been applied prior to use of Thompson and Sharp's test. A s trend towards more rejections for larger sample sizes can be seen.

@@

The results of this can be seen in Figure \ref{significance.change.fixed} for all three methods, comparing the significance of the corrected pooled effect size estimate with the significance of the pooled effect size estimate of the fixed effects meta-analysis. The same for significance of random effects meta-analysis is shown in Figure \ref{significance.change.random}. The significance threshold was chosen such that the $p$-value had to be $< 0.05$ for rejection of the null hypothesis of no treatment effect. The correction methods were trim-and-fill, copas selection model and regression with random effects and shrinkage of within-study-variance methods. More details to the applied correction methods and their application are in the methods section \ref{methods}. Notably, the correction methods has been applied to all meta-analyses, thus also for such that had no significant small study effect test result.

>>>>

<<echo=FALSE>>=
library(knitr)
opts_chunk$set(fig.path='figure/ch02_fig',
               echo=TRUE, message=FALSE,
               fig.width=8.1, fig.height=3.4,
               out.width='\\textwidth-3cm',
               message=FALSE, fig.align='center',
               background="gray98", tidy=FALSE, #tidy.opts=list(width.cutoff=60),
               cache=TRUE
)
options(width=74)
@

@@

\begin{figure}
<<echo=FALSE>>=
levels(data$outcome.measure.new) <- c("Mean Difference", "Std. Mean Difference", "Odds Ratio", "Risk Ratio")

data %>% filter(total1 + total2 > 10 & total1 + total2 < 100) %>% 
  filter(outcome.measure.new == "Risk Ratio" | outcome.measure.new == "Odds Ratio" | outcome.measure.new == "Mean Difference" |
           outcome.measure.new == "Std. Mean Difference") %>% 
  group_by(outcome.measure.new) %>% 
  mutate(sample.size = total1 + total2, scaled.effect = abs(scale(effect, center = T, scale = T))) %>%
  group_by(sample.size, outcome.measure.new) %>% 
  summarize(median.effect = median(scaled.effect, na.rm = T)) %>% 
  ggplot(aes(x = sample.size, y = median.effect, group = outcome.measure.new)) + geom_point() + facet_wrap(~outcome.measure.new, scales = "free") + 
  theme_bw() + xlab("sample size") + ylab("median absolute normalized effect size")
@
\caption{Median of the absolute value of the normalized effect size plotted against the total sample size, separated for outcome measures.}
\label{effect.samplesize.separated}
\end{figure}

% A second way to analyse meta-analyses is a cumulative meta-analysis that should reveal shifts in treatment effect sizes over time. This can again be done for the entire dataset, i.e. for all effect estimates and their ordering in time. It is important to scale the effect estimates here with respect to the estimates of the replication studies that are about the same subject, have the same outcome measure, etc. Also, only effect estimates that can be compared to other estimates before are included, i.e. only study results with one or more replica are included. Time needs to be scaled and normalized in order to compare multiple time-trends in effect size to each other and gain insights in the overall trend. This is done in Figure \ref{effects.overtime.separated}, separared for odds rati, risk ratio, and mean and standardized difference outcome measures. In order to reduce the spread over time of effect sizes, only studies between 1970 and 2019 were included, as well as studies with a minimal group size of 12 participants (each group).
% 
% \begin{figure}
% <<echo=FALSE>>=
% data.ext %>% filter(outcome.measure.new == "Mean Difference" | outcome.measure.new == "Std. Mean Difference" | 
%                 outcome.measure.new == "Risk Ratio" | outcome.measure.new == "Odds Ratio") %>% 
%   filter(total1 > 11 & total2 > 11) %>% filter(!is.na(study.year) & study.year > 1970 & study.year < 2019) %>% 
%   group_by(meta.id) %>% mutate(n = n()) %>% filter(n > 1) %>% 
%   mutate(scaled.effect = abs(scale(effect)), scaled.time = scale(study.year)) %>% 
%   ggplot(aes(x = scaled.time, y = scaled.effect)) + geom_point(size = .5, alpha = 0.2)  + facet_wrap(~outcome.measure.new) +
%   geom_smooth() + theme_bw() +  ylab("absolute scaled effect")  + ylab("scaled time")
% @
% \caption{Absolute values of scaled effect sizes over scaled time, separated for outcome measures.}
% \label{effects.overtime.separated}
% \end{figure}

@@

\vspace{0mm}
Test performance depends on the sample size, despite having restricted sample size to a minimum of 10 studies. The p-values of the Thompson and Sharp tests are shown with respect to the sample size of the meta-analysis in Figure \ref{pvalues.samplesize}. %(Suggests that 10 is likely too small)
In the case of binary outcomes, the arcsine variance stabilizing function has been applied prior to use of Thompson and Sharp's test. A trend towards more rejections for larger sample sizes can be seen.

@@

The results of this can be seen in Figure \ref{significance.change.fixed} for all three methods, comparing the significance of the corrected pooled effect size estimate with the significance of the pooled effect size estimate of the fixed effects meta-analysis. The same for significance of random effects meta-analysis is shown in Figure \ref{significance.change.random}. The significance threshold was chosen such that the $p$-value had to be < 0.05 for rejection of the null hypothesis of no treatment effect. The correction methods were trim-and-fill, copas selection model and regression with random effects and shrinkage of within-study-variance methods. More details to the applied correction methods and their application are in the methods section \ref{methods}. Notably, the correction methods has been applied to all meta-analyses, thus also for such that had no significant small study effect test result.
